---
title: "p8105_hw2_jh4617"
author: "Junjie Hu"
date: "2023-10-01"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

Clean *pols* dataframe

```{r clean pols_month}
pols = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day") )|>
  mutate(president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  mutate(month = recode(month,
                        "01" = "jan",
                        "02" = "feb",
                        "03" = "mar",
                        "04" = "apr",
                        "05" = "may",
                        "06" = "jun",
                        "07" = "jul",
                        "08" = "aug",
                        "09" = "sep",
                        "10" = "oct",
                        "11" = "nov",
                        "12" = "dec")) |>
  select(-prez_dem, -prez_gop, -day) |>
  arrange (year, month)
    
head(pols)
```

Clean *snp* dataframe

```{r clean snp}
snp = 
  read_csv("fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>
  mutate(date = as.Date(date, format = "%m/%d/%y")) |>
  separate(date, into = c("year", "month", "day")) |>
  mutate(year = as.numeric(year), month = as.numeric(month)) |>
  mutate(year = ifelse(year > as.integer(format(Sys.Date(),"%Y")),
                       year - 100, year)) |>
  mutate(month = recode(month,
                        "1" = "jan",
                        "2" = "feb",
                        "3" = "mar",
                        "4" = "apr",
                        "5" = "may",
                        "6" = "jun",
                        "7" = "jul",
                        "8" = "aug",
                        "9" = "sep",
                        "10" = "oct",
                        "11" = "nov",
                        "12" = "dec"))|>
  select(-day) |>
  arrange(year, month) |>
  mutate(year = as.character(year))

head(snp)
tail(snp)


```

Clean *unemployment* dataframe

```{r clean unemployment}
unemployment = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment") |> 
  select(year, month, unemployment) |>
  mutate(year = as.character(year))

head(unemployment)
tail(unemployment)
```

Merge *pols*, *snp*, and *unemployment* dataframes together

```{r}
problem1_df = left_join(pols, snp) |> 
  left_join(unemployment, copy = TRUE)

problem1_df
```




# Problem 2

Clean *Mr.trash* dataframe

```{r mr.trash}
mr.trash = 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 1) |>
  janitor::clean_names() |>
  mutate(homes_powered = weight_tons*500/30) |>
  drop_na(dumpster)

mr.trash
```

Clean *professor_trash* and *gwynnada_trash* dataframe

```{r other trash}
professor_trash = 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 2) |>
  janitor::clean_names() |>
  mutate(homes_powered = weight_tons*500/30) |>
  mutate(year = as.character(year)) |>
  drop_na(dumpster)

professor_trash

gwynnda_trash = 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 4) |>
  janitor::clean_names() |>
  mutate(homes_powered = weight_tons*500/30) |>
  mutate(year = as.character(year)) |>
  drop_na(dumpster)

gwynnda_trash
```

Merage *mr.trash*, *professor_trash*, and *gwynnda_trash* dataframes

```{r merge files}
trash_df =
  full_join(mr.trash, professor_trash) |>
  full_join(gwynnda_trash, copy = TRUE) |>
  select(-x15, -x16)

trash_df
```

There are `r nrow(mr.trash)` observations in the Mr.trash file.

There are `r nrow(professor_trash)` observations in the professor_trash file.

There are `r nrow(gwynnda_trash)` observations in the gwynnda_trash file.

There are `r nrow(trash_df)` observations in merged data, including `r ncol(trash_df)` variables and key variables `r names(trash_df)`. 

The total weight of trash collected by Professor Trash Wheel is `r sum(pull(professor_trash, weight_tons))`. 

The total number of cigarette butts collected by Gwynnda in July of 2021 is 
`r gwynnda_trash_filtered =
  gwynnda_trash |>
  filter(year == "2021", month == "July") 
  sum(pull(gwynnda_trash_filtered, cigarette_butts)) `.

  
# problem 3

Clean *MCI_baseline* data

```{r}
MCI_baseline = 
  read.csv("data_mci/MCI_baseline.csv")|>
  slice(-1) |>
  magrittr::set_colnames(c("ID", "current_age", "sex", "education", "apoe4", "age_at_onset"))|>
  janitor::clean_names() |>
  mutate(age_at_onset = ifelse(age_at_onset == ".", NA, age_at_onset),
         current_age = as.numeric(current_age),
         age_at_onset = as.numeric(age_at_onset),
         sex = recode(sex, "0" = "male", "1" = "female"), 
         apoe4 = recode(apoe4, "0" = "non-carrier", "1" = "carrier")) |>
  filter(current_age < age_at_onset | is.na(age_at_onset))
```

```{r echo = FALSE}
MCI_baseline_onset = 
  MCI_baseline |>
  filter(current_age < age_at_onset)

MCI_baseline_female =
  MCI_baseline |>
  filter(sex == "female")

MCI_baseline_female_apoe =
  MCI_baseline_female |>
  filter(apoe4 == "carrier")
```

There are `r nrow(MCI_baseline)` participants were recruited.

Of the recruited participants, `r nrow(MCI_baseline_onset)` participants have developed MCI.

The average baseline age is `r mean(pull(MCI_baseline, current_age))`.

There are `r nrow(MCI_baseline_female_apoe) / nrow(MCI_baseline_female) *100` % of women in the study are APOE4 carriers.
 
 
Clean *mci_amyloid* datafile
 
```{r}
mci_amyloid = 
  read.csv("data_mci/mci_amyloid.csv") |>
  slice(-1) |>
  magrittr::set_colnames(c("ID","baseline","time2","time4","time6","time8")) |>
  janitor::clean_names() |>
  pivot_longer(
    baseline:time8,
    names_to = "time",
    values_to = "amyloid_ratio" 
  )
  
mci_amyloid
```










